Advanced Reactor Fuel Cycle (ARFC) Molten Salt Reactor Design

### 2016-10-04 MSR-design.tex
spellcheck

The basis for this quick synopsis comes from
\cite{wiki:Molten_salt_reactor}. Research into molten salt reactors (MSRs) began
in earnest with the Aircraft Reactor Experiment (ARE), with experiments shared
between Oak Ridge National Lab (ORNL) and what is now Idaho National Lab
(INL). The ARE used NaF-ZrF$_4$-UF$_4$ (53-41-6 mol\%) as fuel. The reactor was
moderated by beryllium oxide (BeO), used liquid sodium as the secondary coolant,
and had a peak tempeature around 860 \textdegree C. The ARE managed to generate
100 MWh over the course of nine days in 1954. There was another reactor that
went critical at ORNL in 1957 called the Pratt and Whitney Aircraft Reactor-1
(PWAR-1). Although it produced essentially no power, the PWAR-1 is still one of
only three critical MSRs ever built.

@@

A cutaway diagram of the MSRE reactor is shown in \cref{fig:MSRE_reactor}. Fuel
enters through the inlet and is distributed evenlgy along the outside of the
reactor vessel in a small annular region of 1 inch
width. \cite{robertson_msre_1965} The fluid fuel then flows turbulently
downward, cooling the reactor vessel to within 5\textdegree F of the entering
bulk salt temperature (1175 \textdegree F). \cite{robertson_msre_1965} Some MSRE
modeling papers (\cite{kophazi_development_????}) refer to this down-flowing
region as the \textit{downcomer}. The fuel salt is collected in the lower plenum
(lower \textit{space}) of the reactor where straightening vanes remove the
salt's rotational motion. The salt then flows upward through approximately 1140
0.4-inch by 1.2-inch fuel channels that are formed by machinging grooves in the
sides of the graphite moderator columns \cite{robertson_msre_1965}; these are the
regions where fission occurs. The graphite columns themselves have 2-inch by
2-inch square cross sections. Nominal core volume is 90 ft$^3$, with 70 ft$^3$
occupied by graphite and 20 ft$^3$ occupied by fuel. \cite{robertson_msre_1965}
After flowing through the core channels, the fuel collects in the upper plenum
before exiting through the outlet.

>>>>

The basis for this quick synopsis comes from
\cite{wiki:Molten_salt_reactor}. Research into molten salt reactors (MSRs) began
in earnest with the Aircraft Reactor Experiment (ARE), with experiments shared
between Oak Ridge National Lab (ORNL) and what is now Idaho National Lab
(INL). The ARE used NaF-ZrF$_4$-UF$_4$ (53-41-6 mol\%) as fuel. The reactor was
moderated by beryllium oxide (BeO), used liquid sodium as the secondary coolant,
and had a peak temperature around 860 \textdegree C. The ARE managed to generate
100 MWh over the course of nine days in 1954. There was another reactor that
went critical at ORNL in 1957 called the Pratt and Whitney Aircraft Reactor-1
(PWAR-1). Although it produced essentially no power, the PWAR-1 is still one of
only three critical MSRs ever built.

@@

A cutaway diagram of the MSRE reactor is shown in \cref{fig:MSRE_reactor}. Fuel
enters through the inlet and is distributed evenly along the outside of the
reactor vessel in a small annular region of 1 inch
width. \cite{robertson_msre_1965} The fluid fuel then flows turbulently
downward, cooling the reactor vessel to within 5\textdegree F of the entering
bulk salt temperature (1175 \textdegree F). \cite{robertson_msre_1965} Some MSRE
modeling papers (\cite{kophazi_development_????}) refer to this down-flowing
region as the \textit{downcomer}. The fuel salt is collected in the lower plenum
(lower \textit{space}) of the reactor where straightening vanes remove the
salt's rotational motion. The salt then flows upward through approximately 1140
0.4-inch by 1.2-inch fuel channels that are formed by machining grooves in the
sides of the graphite moderator columns \cite{robertson_msre_1965}; these are the
regions where fission occurs. The graphite columns themselves have 2-inch by
2-inch square cross sections. Nominal core volume is 90 ft$^3$, with 70 ft$^3$
occupied by graphite and 20 ft$^3$ occupied by fuel. \cite{robertson_msre_1965}
After flowing through the core channels, the fuel collects in the upper plenum
before exiting through the outlet.
